---
title: "Identify and Trace Plume Emission Source"
description: " "
date: "2025-04-27"
categories: [R, HPAI, Spatiotemporal]
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 3
    toc-location: right
    number-sections: true
    number-depth: 2
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options:  
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```


## Overview
This script provides a conceptual workflow that *back traces* or verifies the source farm that instigated secondary outbreaks at other farms through atmospheric dispersion of virus.  That is, the script looks at data from a secondary outbreak and attempts to determine what source farm emitted the virus that seeded virus introduction.  This script moves *backward* in time to identify the source of a pathogenic plume.      
  
The companion and preceding script for this analysis is available here, [Simulated Outbreak](https://geoepi.github.io/EpiPlume/simulate_outbreak.html).  That analysis moves *forward* in time to model what farms could have been subject to pathogen invasion after plume emission from a source farm. This preliminary script provides a more intuitive overview than the current one which requires a hindsight perspective.  
  
  
## Preliminary Setup

### Libraries
This workflow uses the **renv** package for package version and environmental management.  The `renv.lock` file is included to facilitate package installation. See, *renv::restore()*.
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(gganimate)
library(here)
library(ggmap)
library(ggspatial)
library(yaml)
library(sf)
library(terra)
library(raster)
select <- dplyr::select
library(deSolve)
library(gt)
library(pals)
library(osfr)

library(INLA) # Bayesian spatial temporal modeling
# install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)

library(splitr) # HYSPLIT wrappers
# devtools::install_github("rich-iannone/splitr")
```

### Core Functions
Custom functions
```{r}
source(here("R/utilities.R"))
source_dir(here("R"))
```

### Authentication
User credentials to access data repository and API resources.
```{r}
authentications <- yaml::read_yaml(here("local", "secrets.yaml"))
```
  
### API Registration 
**Stadia** maps are used for backgrounds.  
```{r}
register_stadiamaps(key = authentications$stadi_api)
```
  
### OSF Link
Models and data generated from this script are stored at the *Open Science Framework (OSF)* for remote access.  See, this [**OSF Project:**](https://osf.io/98hmt/)  
```{r}
osf_auth(authentications$osf_token) # for write/admin privileges

osf_project_demo <- osf_retrieve_node("https://osf.io/su4yz/")

osf_project_demo$name # scratch directory on OSF EpiPlume project
```
  
## Read Configuration Parameters  
Parameters for epidemiological and spatial analyses are recorded in configuration files.  Loading the demonstration config file, which can be viewed here: [Demo Configuration Parameters](https://github.com/JMHumphreys/EpiPlume/blob/main/config/demo_2025-04-21.yaml).
```{r}
cfg <- yaml::read_yaml(here("config", "demo_2025-04-21.yaml"))
```
  

## Simulated Outbreak
Downloading and reading the data frame resulting from forward simulation in the Simulated Outbreak script. This file has the coordinates and outbreak detection date for the secondarily infected farms (only one in this demo run) and source farm.   
```{r}
osf_id <- osf_project_demo %>%
  osf_ls_files() %>%
  filter(name == "demo_outbreak_sim_2025-04-28.csv")

osf_download(osf_id,
             path = here("local/temp"),
             conflicts = "overwrite")

simulated_outbreaks <- read.csv(here("local/temp/demo_outbreak_sim_2025-04-28.csv")) %>%
  filter(!is.na(outbreak)) # keeping only farms that experienced outbreak
```

### Convert to Points
Data frame to spatial points
```{r}
# Long/Lat
farm_points_wgs84 <- vect(simulated_outbreaks,
                          geom = c("x", "y"),
                          crs = "EPSG:4326")
```

Get metadata for simulated farms
```{r}
outbreak_farm_meta <- as.data.frame(farm_points_wgs84, geom="xy") %>%
  filter(name != "source_farm") %>%
  select(-X) # row names

orig_source_meta <- as.data.frame(farm_points_wgs84, geom="xy") %>%
  filter(name == "source_farm") %>%
  select(-X) 
```

## Create Study Area
Defining a study area extent (raster grid) with the secondary outbreak farm at center. The original source farm is also depicted as reference. The *define_study_domain()* identifies the appropriate UTM and returns projected objects. 
```{r}
study_area_args <- c(
  list(
    source_origin = c(outbreak_farm_meta$x, outbreak_farm_meta$y),
    extent_km = cfg$domain$extent_km,
    res_m = cfg$domain$res_m  
  )
)

study_area <- exec(define_study_domain, !!!study_area_args)

# projected UTM
farm_points <- project(farm_points_wgs84, crs(study_area$projection))
```
 
### Plot study area
The circle at center is the secondarily infected farm; the goal is to look backward in time to demonstrate that the source farm (triangle, lower right) emitted the plume that gave rise to the outbreak at the center.
```{r warning=FALSE, message=FALSE, fig.width=8, fig.height=8}
map_grid(study_area$grid, source_loc = farm_points)
```

## Trajectory Models
A [HYSPLIT](https://www.ready.noaa.gov/HYSPLIT.php) model is constructed assuming the secondary outbreak location was a *receptor* of the plume from an unknown source.  The [**splitR** package](https://github.com/rich-iannone/splitr) is used to call a HYSPLIT executable from R.
   
Time limit to download meteorological data.
```{r}
getOption("timeout")
options(timeout=10000)
```
   

### Date Determination
Determine dates and times to evaluate. Although it is known when the outbreak was detected at the secondary farm, it is not known when the virus was introduced nor when initial infection occurred. In practice, poultry workers would likely only detect an outbreak once the rate of daily chicken mortality sufficiently and consistently (multiple days) exceeded that typical background rate.  In this example, it's assumed that latent periods and initial spread from a single chicken took at least 3 days (earliest possible detection), but may have taken 20 days to detect. Therefore, this demo analyses more than two weeks in the past.  
```{r}
model_name <- gsub("_", "", outbreak_farm_meta$name) # 6 characters max
model_name

earliest_date <- as_date(outbreak_farm_meta$outbreak) - days(20) # assuming slow to detect
latest_date <- as_date(outbreak_farm_meta$outbreak) - days(3) # allowing for at least 2 day latent period and 1 day infectious

dates_to_check <- seq(earliest_date, latest_date, by = 1)
length(dates_to_check) # total days to model
```

### Run Trajectory Model  
The *wrap_trajectory_model()* saves locally, but a copy is also saved to OSF for others to use. There are a ton of parameter assumptions here to discuss.  Brief explanations are provided with the [config file (linked here)](https://github.com/JMHumphreys/EpiPlume/blob/main/config/demo_2025-04-21.yaml), but its too much for this draft workflow.
   
This model estimates four trajectories per day (every 6 hours) for each of the `r length(dates_to_check)` days, for a a total of `r 4*length(dates_to_check)` trajectories.  
```{r eval=FALSE}
traj_model <- wrap_trajectory_model(cfg, # configuration file
                                    traj_name = model_name,
                                    lon = outbreak_farm_meta$x, # simulated outbreak farm
                                    lat = outbreak_farm_meta$y, # simulated outbreak farm
                                    height = cfg$trajectory$traj_height,
                                    duration = cfg$trajectory$traj_duration,
                                    days = dates_to_check, # days to assess based on simulated outbreak timeline
                                    daily_hours = cfg$trajectory$daily_hours,
                                    model_height = cfg$trajectory$model_height, 
                                    direction = cfg$trajectory$traj_direction,
                                    extended_met = cfg$trajectory$extended_met,
                                    vert_motion = cfg$trajectory$vert_motion,
                                    met_type = cfg$trajectory$traj_met_type,
                                    met_dir = here(cfg$trajectory$traj_climate),
                                    exec_dir = here(cfg$trajectory$traj_outputs),
                                    clean_up = cfg$trajectory$traj_clean_up
                                )

# save copy to OSF
osf_upload(osf_project_demo, path = here("local/demo_run/traj/farm18_model.rds"))
```

Download trajectory model from OSF (if not run locally)
```{r eval=FALSE, message=FALSE, warning=FALSE}
osf_id <- osf_project_demo %>%
  osf_ls_files() %>%
  filter(name == "farm18_model.rds")

osf_download(osf_id,
             path = here("local/temp"),
             conflicts = "overwrite")
```
  
Read the pre-run trajectory model  
```{r}
traj_model <- readRDS(here("local/temp/farm18_model.rds"))
```

### Plot hindcast trajectories 
Each line is a trajectory extending from the secondarily infected farm (center) that follows atmospheric characteristics for the hind cast date and time.  This plot shows the complete trajectory.  
```{r fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
map_grid2(study_area$grid, traj_model$traj_df, group_col = "run", vector_type = "line", grid_cut = FALSE)
```


The trajectories in this plot are cropped to the extent of study area with option `grid_cut` to better see individual paths.  
    
**Question:** Do any of these trajectories help identify and confirm the source farm?   
```{r fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
map_grid3(study_area$grid, traj_model$traj_df, group_col = "run", vector_type = "line", source_loc = farm_points_wgs84, grid_cut = TRUE)
```


## Spatial Analysis
A couple basic steps to compare distances between the source farm and possible trajectories from the secondary outbreak farm.

### Measure distances
The *measure_proximity()* function measures distances for both the trajectory lines and point outputs.  
```{r}
traj_dists <- measure_proximty(
  target_location = c(orig_source_meta$x, orig_source_meta$y), # suspected source farm
  traj_model = traj_model,
  projection = study_area$projection
  )
```

  
### Inverse Distance Decay
Apply an inverse distance decay function assuming that trajectory probability decreases the further from the source farm it is:

$$
w(d) = \exp(-\lambda \times d)
$$  
where:  
- $w(d)$ is the weight (between 0 and 1),  
- $d$ is the distance in meters between the source and the trajectory,  
- $\lambda$ is a decay rate parameter, how quickly the weight decreases with distance.  
     
 
```{r}
distance_threshold <- 10000 # meters, 10k, decreasing probability of trajectory after this distance

lambda_value <- calibrate_lambda(distance_threshold, weight_target = 0.5) # find the best lambda with 10k as the median probability
```
   
Show the relationship (green is the `lambda_value`).   
```{r fig.width=8, fig.height=8}
plot_distance_decay(
    lambda_target = lambda_value,
    weight_target = 0.5 # median probability
)
```
  
### Score and Subset    
Calculate the probability for each modeled trajectory
```{r}
traj_dists$traj_lines$score <- calculate_inverse_distance_weight(traj_dists$traj_lines$distance_m, 
                                                                 lambda = lambda_value) # score distances
```

Select trajectories with at least a 90\% probability.  
```{r}
closest_traj_line <- as.data.frame(traj_dists$traj_lines) %>%
  filter(score >= 0.9) %>%
  arrange(desc(score)) %>%
  select(run, distance_m, score)
```

Filter trajectory data to possible distances.
```{r}
runs_50 <- unique(closest_traj_line$run)

possible_dates <- vect(traj_dists$traj_point)
possible_dates <- project(possible_dates, "EPSG:4326" )
possible_dates <- as.data.frame(possible_dates,
                                geom="xy") %>%
  mutate(lon = x, lat = y) %>%
  filter(run %in% runs_50)
```
   
The above reduced `r 4*length(dates_to_check)` possible trajectories to `r unique(as_date(possible_dates$run))` probable ones.  Show here: 
```{r fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
map_grid3(study_area$grid, possible_dates, group_col = "run", vector_type = "line", source_loc = farm_points_wgs84, grid_cut = TRUE)
```


## Receptor Return
Having narrowed possible trajectory dates and times that could have produced the infecting plume, reverse plumes are simulated from the secondary outbreak farm (receptor) to the original source (backwards in time).
  
  
Execute plume model runs in a loop.
```{r eval=FALSE}
possible_datetimes <- possible_dates %>%
  mutate(date = as.Date(traj_dt)) %>%
  group_by(date) %>%
  summarise(start = min(traj_dt)) %>%
  as.data.frame()

for(i in 1:unique(possible_datetimes$start)){
  
  plume_name <- paste0("plume_", i)
  
  plume_model <- wrap_plume_model(cfg, # configuration file
                                plume_name = plume_name, # name/label
                                lon = outbreak_farm_meta$x, # secondary outbreak farm
                                lat = outbreak_farm_meta$y, # secondary outbreak farm
                                direction = "backward", # from receptor back in time
                                release_start = possible_datetimes$start[i], # release from source farm
                                release_end = possible_datetimes$start[i] - hours(1), # reverse time
                                start_time = possible_datetimes$start[i], # dispersion time same as emission
                                end_time = possible_datetimes$start[i] - hours(24) # dispersion end
                                )
  
  
  
}

# copy to OSF
osf_upload(osf_project_demo, path = here("local/demo_run/plume/plume_1_model.rds"))
osf_upload(osf_project_demo, path = here("local/demo_run/plume/plume_2_model.rds"))
osf_upload(osf_project_demo, path = here("local/demo_run/plume/plume_3_model.rds"))
```
 
Download from OSF
```{r eval=FALSE}
osf_id <- osf_project_demo %>%
  osf_ls_files() %>%
  filter(name == "plume_1_model.rds" | 
         name == "plume_2_model.rds" |
         name == "plume_3_model.rds")

osf_download(osf_id,
             path = here("local/temp"),
             conflicts = "overwrite")
```
  
Read plume models  
```{r}
plume_1 <- readRDS(here("local/temp/plume_1_model.rds"))
plume_2 <- readRDS(here("local/temp/plume_2_model.rds"))
plume_3 <- readRDS(here("local/temp/plume_3_model.rds"))
```

### Plume 1 simulated dispersion
```{r fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
map_grid3(study_area$grid, plume_1$disp_df, group_col = "particle_i", vector_type = "point", source_loc = farm_points_wgs84, grid_cut = TRUE)
```

### Plume 2 simulated dispersion
```{r fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
map_grid3(study_area$grid, plume_2$disp_df, group_col = "particle_i", vector_type = "point", source_loc = farm_points_wgs84, grid_cut = TRUE)
```

### Plume 3 simulated dispersion
```{r fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
map_grid3(study_area$grid, plume_3$disp_df, group_col = "particle_i", vector_type = "point", source_loc = farm_points_wgs84, grid_cut = TRUE)
```


