---
title: "User-configurable EpiPlume demonstration diagnostic"
format:
  html:
    toc: true
    embed-resources: true
    html-math-method: plain
params:
  run_root: null
execute:
  echo: false
  warning: false
---

```{r}
audit_source <- if (file.exists("R/audit_demo_run.R")) "R/audit_demo_run.R" else "../R/audit_demo_run.R"
source(audit_source)
root <- params$run_root
`%||%` <- function(x, y) if (is.null(x) || !length(x)) y else x
if (is.null(root) || !dir.exists(root)) stop("A prepared `run_root` parameter is required.")
root <- normalizePath(root, winslash = "/", mustWork = TRUE)
read_csv <- function(path) if (file.exists(path)) utils::read.csv(path, stringsAsFactors = FALSE) else data.frame()
read_yml <- function(path) if (file.exists(path) && requireNamespace("yaml", quietly = TRUE)) yaml::read_yaml(path) else list()
column_or <- function(x, name, default = NA) if (name %in% names(x)) x[[name]] else rep(default, nrow(x))
pretty_number <- function(x, digits = 0L) format(round(as.numeric(x), digits), big.mark = ",", scientific = FALSE, trim = TRUE, nsmall = digits)
report_table <- function(x, col.names = names(x), digits = 1L) knitr::kable(x, format = "html", col.names = col.names, digits = digits, format.args = list(big.mark = ","), table.attr = 'class="table table-striped table-hover table-sm"')
select_for_report <- function(x, labels) {
  fields <- intersect(names(labels), names(x))
  out <- x[, fields, drop = FALSE]
  names(out) <- unname(labels[fields])
  out
}
finite_summary <- function(x) {
  x <- suppressWarnings(as.numeric(x)); x <- x[is.finite(x)]
  if (!length(x)) return(c(n = 0, median = NA, minimum = NA, maximum = NA))
  c(n = length(x), median = stats::median(x), minimum = min(x), maximum = max(x))
}
facilities <- read_csv(file.path(root, "inputs", "facilities.csv"))
schedule <- read_csv(file.path(root, "inputs", "release_schedule.csv"))
manifest <- read_csv(file.path(root, "inputs", "run_manifest.csv"))
met <- read_csv(file.path(root, "meteorology", "meteorology_inventory.csv"))
audit <- read_csv(file.path(root, "combined", "run_audit.csv"))
status <- read_csv(file.path(root, "combined", "run_status_summary.csv"))
shards <- read_csv(file.path(root, "manifests", "shard_manifest.csv"))
history <- inventory_demo_shard_history(root)
prov <- read_yml(file.path(root, "provenance", "preparation_provenance.yml"))
cfg <- read_yml(file.path(root, "inputs", "config.yml"))
exchange_files <- list.files(file.path(root, "runs"), pattern = "source_receptor_exchange[.]csv$", recursive = TRUE, full.names = TRUE)
exchange <- if (length(exchange_files)) do.call(rbind, lapply(exchange_files, read_csv)) else data.frame()
intercept <- if ("intercept" %in% names(exchange)) as.logical(exchange$intercept) else if ("binary_intercept" %in% names(exchange)) as.logical(exchange$binary_intercept) else rep(NA, nrow(exchange))
n_intercepts <- sum(intercept %in% TRUE)
source_ids <- if ("source_facility_id" %in% names(manifest)) unique(as.character(manifest$source_facility_id)) else if ("source_id" %in% names(manifest)) unique(as.character(manifest$source_id)) else character()
coverage_ok <- nrow(manifest) == nrow(audit) && nrow(audit) > 0L && !anyDuplicated(audit$run_id) && setequal(manifest$run_id, audit$run_id)
diagnostic <- demo_diagnostic_result(audit, coverage_ok)
n_completed <- sum(column_or(audit, "execution_status", "unknown") == "completed_valid")
parsed_total <- sum(column_or(audit, "parsed_row_count", NA_real_), na.rm = TRUE)
parsed_available <- any(!is.na(column_or(audit, "parsed_row_count", NA_real_)))
receptor_total <- if (nrow(exchange)) nrow(exchange) else sum(column_or(audit, "receptor_row_count", NA_real_), na.rm = TRUE)
receptor_available <- nrow(exchange) > 0L || any(!is.na(column_or(audit, "receptor_row_count", NA_real_)))
release_values <- if ("release_datetime_utc" %in% names(manifest)) manifest$release_datetime_utc else if ("release_start" %in% names(manifest)) manifest$release_start else character()
release_times <- as.POSIXct(release_values, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
release_times <- release_times[!is.na(release_times)]
release_text <- if (!length(release_times)) "an unavailable release-time range" else if (min(release_times) == max(release_times)) format(min(release_times), "%B %d, %Y at %H:%M UTC", tz = "UTC") else paste(format(min(release_times), "%B %d, %Y at %H:%M UTC", tz = "UTC"), "through", format(max(release_times), "%B %d, %Y at %H:%M UTC", tz = "UTC"))
duration_values <- if ("duration_hours" %in% names(manifest)) unique(manifest$duration_hours) else cfg$simulation$duration_hours %||% NA_real_
duration_values <- suppressWarnings(as.numeric(duration_values)); duration_values <- duration_values[is.finite(duration_values)]
duration_text <- if (!length(duration_values)) "not available" else if (length(unique(duration_values)) == 1L) paste0(pretty_number(duration_values[1]), " hours") else paste0(pretty_number(min(duration_values)), "-", pretty_number(max(duration_values)), " hours")
```

Run: `r basename(root)` | Prepared: `r prov$prepared_at_utc %||% "unavailable"` | Preparation commit: `r prov$git_commit_sha %||% "unavailable"`

## Executive summary

```{r}
design_sentence <- sprintf("This demonstration evaluated %s atmospheric dispersion simulation%s from %s source facilit%s within a network of %s facilities. Releases span %s, and simulation duration was %s.", pretty_number(nrow(manifest)), if (nrow(manifest) == 1L) "" else "s", pretty_number(length(source_ids)), if (length(source_ids) == 1L) "y" else "ies", pretty_number(nrow(facilities)), release_text, duration_text)
completion_sentence <- if (nrow(audit) && n_completed == nrow(manifest) && coverage_ok) sprintf("All %s requested runs completed and validated successfully.", pretty_number(n_completed)) else sprintf("%s of %s requested runs are currently completed and validated; %s", pretty_number(n_completed), pretty_number(nrow(manifest)), diagnostic)
output_bits <- character()
if (parsed_available) output_bits <- c(output_bits, sprintf("%s standardized dispersion records", pretty_number(parsed_total)))
if (receptor_available) output_bits <- c(output_bits, sprintf("%s directed run-receptor summaries", pretty_number(receptor_total)))
if (nrow(exchange)) output_bits <- c(output_bits, sprintf("%s met the configured binary intercept criterion", pretty_number(n_intercepts)))
output_sentence <- if (length(output_bits)) paste0("The workflow produced ", paste(output_bits, collapse = ", "), ".") else "Dispersion and receptor outputs are not yet available."
coverage_sentence <- if (coverage_ok) "Manifest coverage is complete and internally consistent." else "Manifest coverage requires attention."
cat(paste(design_sentence, completion_sentence, output_sentence, coverage_sentence))
```

## Study and test configuration

The design summary below describes what was requested. Detailed normalized configuration is retained in the technical section.

```{r}
#| label: tbl-test-design
#| tbl-cap: "Test design. Counts and timing are derived from the normalized facility inventory and immutable run manifest."
design <- data.frame(
  Characteristic = c("Facilities", "Source facilities", "Simulations", "Release period", "Simulation duration", "Manifest coverage"),
  Value = c(pretty_number(nrow(facilities)), pretty_number(length(source_ids)), pretty_number(nrow(manifest)), release_text, duration_text, if (coverage_ok) "Valid" else "Requires attention"),
  check.names = FALSE
)
report_table(design, c("Characteristic", "Value"), digits = 0)
```

## Geographic orientation

The orientation figure is generated entirely from facility coordinates and the offline state-boundary data bundled with the installed `maps` package. If those boundaries are unavailable, the facility network and study extent are still shown without attempting a download.

```{r}
#| label: fig-regional-orientation
#| fig-cap: "Regional and local orientation of the facility network. Red triangles identify atmospheric release sources; blue circles identify receptor-only facilities. Panel A places the network bounding box within state boundaries, while Panel B labels the local facilities. The arrow indicates north and the bar gives an approximate geographic scale."
#| fig-width: 11
#| fig-height: 5.5
draw_north_scale <- function(xlim, ylim, latitude) {
  dx <- diff(xlim); dy <- diff(ylim)
  x_n <- xlim[2] - 0.08 * dx; y0 <- ylim[1] + 0.72 * dy; y1 <- ylim[1] + 0.91 * dy
  arrows(x_n, y0, x_n, y1, length = 0.08, lwd = 1.5); text(x_n, y1, "N", pos = 3, font = 2)
  target_km <- if (dx * 111 < 50) 5 else if (dx * 111 < 150) 10 else 25
  degrees <- target_km / (111.32 * max(cos(latitude * pi / 180), 0.2))
  sx0 <- xlim[1] + 0.08 * dx; sx1 <- sx0 + degrees; sy <- ylim[1] + 0.08 * dy
  segments(sx0, sy, sx1, sy, lwd = 3); segments(c(sx0, sx1), sy - 0.015 * dy, c(sx0, sx1), sy + 0.015 * dy, lwd = 1.5); text(mean(c(sx0, sx1)), sy, paste(target_km, "km"), pos = 3, cex = 0.8)
}
draw_offline_boundaries <- function(xlim, ylim, color) {
  boundary <- if (requireNamespace("maps", quietly = TRUE)) try(maps::map("state", xlim = xlim, ylim = ylim, plot = FALSE), silent = TRUE) else NULL
  if (inherits(boundary, "map")) maps::map(boundary, xlim = xlim, ylim = ylim, col = color, fill = FALSE, mar = c(0, 0, 0, 0)) else plot(NA, xlim = xlim, ylim = ylim, xlab = "Longitude", ylab = "Latitude", asp = 1)
}
if (nrow(facilities) && all(c("longitude", "latitude", "facility_id") %in% names(facilities))) {
  lon <- facilities$longitude; lat <- facilities$latitude; is_source <- facilities$facility_id %in% source_ids
  local_pad_x <- max(diff(range(lon)) * 0.15, 0.03); local_pad_y <- max(diff(range(lat)) * 0.15, 0.03)
  local_x <- range(lon) + c(-local_pad_x, local_pad_x); local_y <- range(lat) + c(-local_pad_y, local_pad_y)
  regional_x <- range(lon) + c(-max(2, diff(range(lon)) * 5), max(2, diff(range(lon)) * 5)); regional_y <- range(lat) + c(-max(1.5, diff(range(lat)) * 5), max(1.5, diff(range(lat)) * 5))
  old <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
  draw_offline_boundaries(regional_x, regional_y, "#8a8a8a")
  rect(local_x[1], local_y[1], local_x[2], local_y[2], border = "#d73027", lwd = 2, lty = 2)
  points(lon, lat, pch = ifelse(is_source, 17, 21), bg = ifelse(is_source, "#d73027", "#74add1"), cex = 0.8)
  title("A. Regional context"); box(); draw_north_scale(regional_x, regional_y, mean(lat))
  draw_offline_boundaries(local_x, local_y, "#c0c0c0")
  points(lon, lat, pch = ifelse(is_source, 17, 21), bg = ifelse(is_source, "#d73027", "#74add1"), col = "#333333", cex = 1.1)
  text(lon, lat, labels = facilities$facility_id, pos = 3, cex = 0.65)
  title("B. Facility network"); box(); draw_north_scale(local_x, local_y, mean(lat))
  legend("bottomright", legend = c("Release source", "Receptor-only"), pch = c(17, 21), pt.bg = c("#d73027", "#74add1"), bty = "n", cex = 0.8)
  par(old)
} else {
  plot.new(); text(0.5, 0.5, "Facility coordinates are unavailable.")
}
```

## Simulation and execution summary

`r if (nrow(audit)) sprintf("%s of %s simulations are completed and validated. %s", pretty_number(n_completed), pretty_number(nrow(manifest)), diagnostic) else "Execution status is unavailable."` Meteorology is `r if (nrow(met) && all(met$ready %in% TRUE)) "ready for all required files" else if (nrow(met)) "incomplete or unavailable for at least one required file" else "not inventoried"`.

```{r}
#| label: tbl-run-status
#| tbl-cap: "Run status. Zero-count categories are omitted for readability; the counts sum to the requested simulations represented in the current audit."
if (nrow(status)) {
  display <- status[status$run_count > 0, , drop = FALSE]
  display$execution_status <- gsub("_", " ", display$execution_status); display$execution_status <- tools::toTitleCase(display$execution_status)
  names(display) <- c("Run status", "Runs")
  report_table(display, names(display), digits = 0)
} else cat("Run-status details are unavailable.")
```

```{r}
#| label: tbl-runtime
#| tbl-cap: "Runtime telemetry by processing stage. HYSPLIT execution and resumed post-processing are kept separate; unavailable telemetry is displayed as unavailable rather than zero."
if (nrow(audit)) {
  execution_runtime <- finite_summary(column_or(audit, "execution_elapsed_seconds", NA_real_))
  resume_runtime <- finite_summary(column_or(audit, "postprocessing_elapsed_seconds", NA_real_))
  runtime <- data.frame(Stage = c("HYSPLIT execution attempts", "Resumed post-processing"), `Runs with telemetry` = c(execution_runtime["n"], resume_runtime["n"]), `Median seconds` = c(execution_runtime["median"], resume_runtime["median"]), `Minimum seconds` = c(execution_runtime["minimum"], resume_runtime["minimum"]), `Maximum seconds` = c(execution_runtime["maximum"], resume_runtime["maximum"]), check.names = FALSE)
  report_table(runtime, names(runtime), digits = 1)
} else cat("Runtime telemetry is unavailable.")
```

## Atmospheric plume diagnostics

The workflow produced `r if (parsed_available) pretty_number(parsed_total) else "an unavailable number of"` standardized dispersion records. The plots below use saved `hourly_particle_count.tif` products only; HYSPLIT is never rerun. At most four completed runs are selected deterministically in manifest order, and each plot shows the saved hourly layer with the largest total particle count.

```{r}
#| label: fig-representative-plumes
#| fig-cap: "Representative saved hourly particle-count rasters for up to the first four completed runs in manifest order. Each panel shows the hour with the largest raster-wide particle total for that run. Red triangles mark release sources and blue circles mark facility receptors; cells without color contain zero or unavailable saved particle counts according to the raster product."
#| fig-width: 8
#| fig-height: 6
completed_ids <- if (nrow(audit)) audit$run_id[audit$execution_status == "completed_valid"] else character()
selected <- head(manifest[manifest$run_id %in% completed_ids, , drop = FALSE], 4L)
plotted <- 0L
if (nrow(selected) && requireNamespace("terra", quietly = TRUE)) {
  for (i in seq_len(nrow(selected))) {
    raster_path <- file.path(selected$run_directory[i], "parsed", "hourly_particle_count.tif")
    if (!file.exists(raster_path)) next
    r <- tryCatch(terra::rast(raster_path), error = function(e) NULL); if (is.null(r) || terra::nlyr(r) < 1L) next
    totals <- tryCatch(as.numeric(terra::global(r, "sum", na.rm = TRUE)[, 1]), error = function(e) rep(NA_real_, terra::nlyr(r)))
    layer <- if (any(is.finite(totals))) which.max(replace(totals, !is.finite(totals), -Inf)) else 1L
    layer_meta <- read_csv(file.path(selected$run_directory[i], "parsed", "raster_layer_metadata.csv")); layer_meta <- layer_meta[layer_meta$value == "particle_count", , drop = FALSE]
    hour <- if (nrow(layer_meta) >= layer) layer_meta$elapsed_hour_bin[layer] else layer - 1L
    terra::plot(r[[layer]], main = paste0(selected$run_id[i], " - particle count, hour ", hour), axes = TRUE)
    if (nrow(facilities) && all(c("longitude", "latitude") %in% names(facilities))) {
      pts <- tryCatch(terra::project(terra::vect(facilities, geom = c("longitude", "latitude"), crs = "EPSG:4326"), terra::crs(r)), error = function(e) NULL)
      if (!is.null(pts)) terra::points(pts, pch = 21, bg = "#74add1", col = "#333333", cex = 0.8)
    }
    source_id <- if ("source_facility_id" %in% names(selected)) as.character(selected$source_facility_id[i]) else as.character(selected$source_id[i]); source <- facilities[facilities$facility_id == source_id, , drop = FALSE]
    if (nrow(source)) {
      src <- tryCatch(terra::project(terra::vect(source, geom = c("longitude", "latitude"), crs = "EPSG:4326"), terra::crs(r)), error = function(e) NULL)
      if (!is.null(src)) terra::points(src, pch = 17, col = "#d73027", cex = 1.4)
    }
    plotted <- plotted + 1L
  }
}
if (!plotted) cat("No completed saved particle-count rasters are available for plotting.")
```

## Source-receptor exchange results

```{r}
if (!nrow(exchange)) cat("Receptor outputs are incomplete or absent. Missing output is not interpreted as zero exposure.") else cat(sprintf("The completed receptor products contain %s directed run-receptor records. %s met the configured binary intercept criterion.", pretty_number(nrow(exchange)), pretty_number(n_intercepts)))
```

```{r}
#| label: tbl-exposure-outcomes
#| tbl-cap: "Source-receptor outcome categories. Counts represent directed run-receptor records after excluding the source facility itself. Outside evaluation distance identifies pairs beyond the configured distance threshold; no modeled exposure identifies valid receptor output with zero modeled particle exposure."
if (nrow(exchange) && "intercept_reason" %in% names(exchange)) {
  outcomes <- as.data.frame(table(exchange$intercept_reason, useNA = "ifany"), stringsAsFactors = FALSE)
  names(outcomes) <- c("Outcome category", "Directed records"); outcomes[[1]] <- tools::toTitleCase(gsub("_", " ", outcomes[[1]]))
  outcomes <- rbind(outcomes, data.frame(`Outcome category` = "Total", `Directed records` = sum(outcomes[[2]]), check.names = FALSE))
  report_table(outcomes, names(outcomes), digits = 0)
} else cat("Exposure outcome categories are unavailable.")
```

```{r}
#| label: tbl-continuous-exposure
#| tbl-cap: "Continuous particle-count exposure summaries. Available, zero, and nonzero directed records remain distinct; screened or unavailable values are not converted to zero."
continuous <- intersect(c("particle_count_max", "particle_count_cumulative"), names(exchange))
if (nrow(exchange) && length(continuous)) {
  summaries <- do.call(rbind, lapply(continuous, function(metric) {
    x <- suppressWarnings(as.numeric(exchange[[metric]])); available <- is.finite(x)
    data.frame(Metric = tools::toTitleCase(gsub("_", " ", metric)), `Available records` = sum(available), `Zero records` = sum(available & x == 0), `Nonzero records` = sum(available & x > 0), Minimum = if (any(available)) min(x[available]) else NA_real_, Median = if (any(available)) stats::median(x[available]) else NA_real_, Maximum = if (any(available)) max(x[available]) else NA_real_, check.names = FALSE)
  }))
  report_table(summaries, names(summaries), digits = 2)
} else cat("Continuous particle-count summaries are unavailable.")
```

```{r}
#| label: tbl-source-summary
#| tbl-cap: "Source summary for positive binary intercepts. The table is suppressed when no source-receptor pair meets the configured criterion."
if (n_intercepts > 0L && all(c("source_id", "receptor_id") %in% names(exchange))) {
  source_summary <- as.data.frame(table(exchange$source_id[intercept %in% TRUE]), stringsAsFactors = FALSE); names(source_summary) <- c("Source facility", "Positive intercepts")
  report_table(source_summary, names(source_summary), digits = 0)
} else cat("No source-level positive-intercept table is shown because no pair met the criterion.")
```

```{r}
#| label: tbl-receptor-summary
#| tbl-cap: "Receptor summary for positive binary intercepts. Counts indicate how often each receptor met the binary criterion across completed source-release simulations."
if (n_intercepts > 0L && all(c("source_id", "receptor_id") %in% names(exchange))) {
  receptor_summary <- as.data.frame(table(exchange$receptor_id[intercept %in% TRUE]), stringsAsFactors = FALSE); names(receptor_summary) <- c("Receptor facility", "Positive intercepts")
  report_table(receptor_summary, names(receptor_summary), digits = 0)
} else cat("No receptor-level positive-intercept table is shown because no pair met the criterion.")
```

```{r}
#| label: fig-positive-links
#| fig-cap: "Facility context and positive binary source-receptor links. Red line segments are drawn only for pairs meeting the configured criterion; source facilities are red triangles and other facilities are blue circles."
#| fig-width: 7
#| fig-height: 6
if (n_intercepts > 0L && all(c("source_longitude", "source_latitude", "receptor_longitude", "receptor_latitude") %in% names(exchange))) {
  is_source <- facilities$facility_id %in% source_ids
  plot(facilities$longitude, facilities$latitude, pch = ifelse(is_source, 17, 21), bg = ifelse(is_source, "#d73027", "#74add1"), col = "#333333", xlab = "Longitude", ylab = "Latitude", main = "Positive binary source-receptor links")
  hit <- which(intercept %in% TRUE); segments(exchange$source_longitude[hit], exchange$source_latitude[hit], exchange$receptor_longitude[hit], exchange$receptor_latitude[hit], col = "#d73027", lwd = 2)
} else {
  plot.new(); text(0.5, 0.5, "No source-receptor pair met the binary intercept criterion.")
}
```

## Key findings

```{r}
findings <- character()
findings <- c(findings, if (nrow(audit) && n_completed == nrow(manifest)) sprintf("All %s requested simulations completed and validated.", pretty_number(n_completed)) else sprintf("%s of %s requested simulations are currently completed and validated.", pretty_number(n_completed), pretty_number(nrow(manifest))))
if (nrow(exchange)) {
  findings <- c(findings, if (n_intercepts > 0L) sprintf("%s of %s directed run-receptor records met the binary intercept criterion.", pretty_number(n_intercepts), pretty_number(nrow(exchange))) else "No directed run-receptor record met the binary intercept criterion.")
  if ("intercept_reason" %in% names(exchange)) {
    outcome_counts <- sort(table(exchange$intercept_reason), decreasing = TRUE); dominant <- tools::toTitleCase(gsub("_", " ", names(outcome_counts)[1])); findings <- c(findings, sprintf("The most common receptor outcome was %s (%s records).", tolower(dominant), pretty_number(outcome_counts[1])))
    outside <- sum(exchange$intercept_reason == "outside_evaluation_distance", na.rm = TRUE); findings <- c(findings, sprintf("%s of %s records were outside the configured evaluation distance.", pretty_number(outside), pretty_number(nrow(exchange))))
  }
  if ("particle_count_cumulative" %in% names(exchange)) {
    value <- suppressWarnings(as.numeric(exchange$particle_count_cumulative)); below <- sum(is.finite(value) & value > 0 & !(intercept %in% TRUE)); findings <- c(findings, sprintf("%s records had nonzero cumulative particle exposure without meeting the binary criterion.", pretty_number(below)))
  }
}
cat(paste0("- ", findings, collapse = "\n"))
```

These are descriptive atmospheric-dispersion diagnostics. They do not establish causal or risk conclusions.

## Technical diagnostics and provenance

Preparation and later execution history are retained separately so retries and post-processing resumes remain transparent.

```{r}
#| label: tbl-execution-history
#| tbl-cap: "Durable SLURM-array attempt history. Execution and resume actions, repository commits, attempt counts, elapsed time, and final states are reported separately when shard telemetry is available."
if (nrow(history)) {
  display <- select_for_report(history, c(submission_id = "Submission", run_id = "Run", action = "Action", repository_commit = "Repository commit", execution_attempted = "Executed HYSPLIT", hysplit_attempt_count_before = "Attempts before", hysplit_attempt_count_after = "Attempts after", elapsed_seconds = "Elapsed seconds", post_state = "Final state", task_exit_status = "Exit status"))
  report_table(display, names(display), digits = 1)
} else cat("No durable SLURM-array attempt history is available.")
```

```{r}
#| label: tbl-incomplete-runs
#| tbl-cap: "Incomplete or failed runs requiring review. The table is omitted when every requested run is completed and validated."
incomplete <- if (nrow(audit)) audit[audit$execution_status != "completed_valid", , drop = FALSE] else data.frame()
if (nrow(incomplete)) {
  incomplete <- select_for_report(incomplete, c(run_id = "Run", shard_id = "Shard", execution_status = "Status", attempt_count = "HYSPLIT attempts", error_message = "Diagnostic message"))
  report_table(incomplete, names(incomplete), digits = 0)
} else cat("No incomplete runs.")
```

```{r}
#| label: tbl-meteorology
#| tbl-cap: "Meteorology inventory and readiness recorded for the prepared run."
if (nrow(met)) {
  met_display <- select_for_report(met, c(required_filename = "Required file", required_by_n_runs = "Runs requiring file", exists = "Exists", size_bytes = "Bytes", ready = "Ready"))
  report_table(met_display, names(met_display), digits = 0)
} else cat("Meteorology inventory is unavailable.")
```

<details>
<summary>Normalized configuration</summary>

```{r}
cat(paste(capture.output(str(cfg, max.level = 3)), collapse = "\n"))
```

</details>

Manifest coverage: **`r if (coverage_ok) "PASS" else "FAIL"`**. Overall diagnostic: **`r diagnostic`**.

## Key output paths

```{r}
#| label: tbl-output-paths
#| tbl-cap: "Key durable artifacts. Absolute paths are retained for Atlas provenance and debugging and intentionally appear only at the end of the report."
paths <- data.frame(
  Artifact = c("Run manifest", "Run audit", "Run-status summary", "Meteorology inventory", "Receptor exchange directory", "Report directory"),
  Path = c(file.path(root, "inputs", "run_manifest.csv"), file.path(root, "combined", "run_audit.csv"), file.path(root, "combined", "run_status_summary.csv"), file.path(root, "meteorology", "meteorology_inventory.csv"), file.path(root, "runs"), file.path(root, "reports")),
  check.names = FALSE
)
report_table(paths, names(paths), digits = 0)
```
