---
title: "User-configurable EpiPlume demonstration diagnostic"
format:
  html:
    toc: true
    embed-resources: true
    html-math-method: plain
params:
  run_root: null
execute:
  echo: false
  warning: false
---

```{r}
audit_source <- if (file.exists("R/audit_demo_run.R")) "R/audit_demo_run.R" else "../R/audit_demo_run.R"
source(audit_source)
root <- params$run_root
`%||%` <- function(x, y) if (is.null(x) || !length(x)) y else x
if (is.null(root) || !dir.exists(root)) stop("A prepared `run_root` parameter is required.")
root <- normalizePath(root, winslash = "/", mustWork = TRUE)
read_csv <- function(path) if (file.exists(path)) utils::read.csv(path, stringsAsFactors = FALSE) else data.frame()
read_yml <- function(path) if (file.exists(path) && requireNamespace("yaml", quietly = TRUE)) yaml::read_yaml(path) else list()
finite_summary <- function(x) {
  x <- suppressWarnings(as.numeric(x)); x <- x[is.finite(x)]
  if (!length(x)) return("Runtime telemetry unavailable")
  sprintf("n=%d; median=%.1f s; min=%.1f s; max=%.1f s", length(x), stats::median(x), min(x), max(x))
}
facilities <- read_csv(file.path(root, "inputs", "facilities.csv"))
schedule <- read_csv(file.path(root, "inputs", "release_schedule.csv"))
manifest <- read_csv(file.path(root, "inputs", "run_manifest.csv"))
met <- read_csv(file.path(root, "meteorology", "meteorology_inventory.csv"))
audit <- read_csv(file.path(root, "combined", "run_audit.csv"))
status <- read_csv(file.path(root, "combined", "run_status_summary.csv"))
shards <- read_csv(file.path(root, "manifests", "shard_manifest.csv"))
history <- inventory_demo_shard_history(root)
prov <- read_yml(file.path(root, "provenance", "preparation_provenance.yml"))
cfg <- read_yml(file.path(root, "inputs", "config.yml"))
coverage_ok <- nrow(manifest) == nrow(audit) && nrow(audit) > 0L && !anyDuplicated(audit$run_id) && setequal(manifest$run_id, audit$run_id)
diagnostic <- demo_diagnostic_result(audit, coverage_ok)
```

## Run and preparation provenance

Run: `r basename(root)`<br>
EpiPlume version: `r prov$epiplume_version %||% "unavailable"`<br>
Preparation commit: `r prov$git_commit_sha %||% "unavailable"`<br>
Prepared UTC: `r prov$prepared_at_utc %||% "unavailable"`

## Execution history

Preparation provenance is retained separately from later execution and post-processing attempts.

```{r}
if (!nrow(history)) {
  cat("No durable SLURM-array attempt history is available.")
} else {
  display <- history[, intersect(c("submission_id", "run_id", "array_job_id", "action", "repository_commit", "execution_attempted", "hysplit_attempt_count_before", "hysplit_attempt_count_after", "elapsed_seconds", "post_state", "task_exit_status"), names(history)), drop = FALSE]
  knitr::kable(display)
  cat("\n\nHYSPLIT attempts by run are reported in the audit below. A `resume_postprocessing` action with `execution_attempted = FALSE` documents post-processing without another atmospheric-model execution.")
}
```

## Normalized configuration

```{r}
cat(paste(capture.output(str(cfg, max.level = 3)), collapse = "\n"))
```

## Facilities and release schedule

There are `r nrow(facilities)` receptor facilities, `r length(unique(schedule$source_facility_id))` source facilities, and `r nrow(schedule)` requested releases.

```{r}
if (nrow(facilities) && all(c("longitude", "latitude", "facility_id") %in% names(facilities))) {
  source_ids <- if ("source_facility_id" %in% names(manifest)) unique(as.character(manifest$source_facility_id)) else if ("source_id" %in% names(manifest)) unique(as.character(manifest$source_id)) else character()
  is_source <- facilities$facility_id %in% source_ids
  plot(facilities$longitude, facilities$latitude, pch = ifelse(is_source, 17, 21), bg = ifelse(is_source, "#d73027", "#74add1"), col = "#333333", cex = 1.2, xlab = "Longitude", ylab = "Latitude", main = "Facilities and atmospheric release sources")
  text(facilities$longitude, facilities$latitude, labels = facilities$facility_id, pos = 3, cex = 0.75)
  legend("topright", legend = c("Release source", "Receptor-only facility"), pch = c(17, 21), pt.bg = c("#d73027", "#74add1"), bty = "n")
} else cat("Facility coordinates are unavailable.")
```

## Manifest and meteorology

The immutable manifest contains `r nrow(manifest)` runs and `r if (nrow(shards)) length(unique(shards$shard_id)) else 0L` shards.

```{r}
if (nrow(met)) knitr::kable(met[, intersect(c("required_filename", "required_by_n_runs", "exists", "size_bytes", "ready"), names(met)), drop = FALSE]) else cat("Meteorology inventory is unavailable.")
```

## Run status

```{r}
if (nrow(status)) knitr::kable(status) else cat("No status summary is available.")
```

## Failed and incomplete runs

```{r}
incomplete <- if (nrow(audit)) audit[audit$execution_status != "completed_valid", intersect(c("run_id", "shard_id", "execution_status", "attempt_count", "error_message"), names(audit)), drop = FALSE] else data.frame()
if (nrow(incomplete)) knitr::kable(incomplete) else cat("No incomplete runs.")
```

## Execution and parsing diagnostics

```{r}
if (nrow(audit)) {
  cat("Completed and validated runs:", sum(audit$execution_status == "completed_valid"), "of", nrow(audit), "\n\n")
  cat("HYSPLIT execution-attempt runtime:", finite_summary(audit$execution_elapsed_seconds), "\n\n")
  cat("Resumed post-processing runtime:", finite_summary(audit$postprocessing_elapsed_seconds), "\n\n")
  cat("Raw output bytes:", sum(audit$raw_output_bytes, na.rm = TRUE), "\n\n")
  parsed_available <- audit$parsed_row_count[!is.na(audit$parsed_row_count)]
  if (length(parsed_available)) cat("Standardized dispersion records:", sum(parsed_available), "across", length(parsed_available), "parsed runs\n\n") else cat("Standardized dispersion record count unavailable.\n\n")
  receptor_available <- audit$receptor_row_count[!is.na(audit$receptor_row_count)]
  if (length(receptor_available)) cat("Source-receptor records:", sum(receptor_available), "across", length(receptor_available), "runs\n") else cat("Receptor output counts unavailable.\n")
} else cat("Execution audit is unavailable.")
```

## Representative particle-count plume diagnostics

Plots use saved `hourly_particle_count.tif` products only; HYSPLIT is never rerun. At most four completed runs are selected deterministically in manifest order. For each run, the displayed hour is the saved raster layer with the largest total particle count.

```{r}
completed_ids <- if (nrow(audit)) audit$run_id[audit$execution_status == "completed_valid"] else character()
selected <- head(manifest[manifest$run_id %in% completed_ids, , drop = FALSE], 4L)
plotted <- 0L
if (nrow(selected) && requireNamespace("terra", quietly = TRUE)) {
  for (i in seq_len(nrow(selected))) {
    raster_path <- file.path(selected$run_directory[i], "parsed", "hourly_particle_count.tif")
    if (!file.exists(raster_path)) next
    r <- tryCatch(terra::rast(raster_path), error = function(e) NULL); if (is.null(r) || terra::nlyr(r) < 1L) next
    totals <- tryCatch(as.numeric(terra::global(r, "sum", na.rm = TRUE)[, 1]), error = function(e) rep(NA_real_, terra::nlyr(r)))
    layer <- if (any(is.finite(totals))) which.max(replace(totals, !is.finite(totals), -Inf)) else 1L
    layer_meta <- read_csv(file.path(selected$run_directory[i], "parsed", "raster_layer_metadata.csv")); layer_meta <- layer_meta[layer_meta$value == "particle_count", , drop = FALSE]
    hour <- if (nrow(layer_meta) >= layer) layer_meta$elapsed_hour_bin[layer] else layer - 1L
    terra::plot(r[[layer]], main = paste0(selected$run_id[i], " — particle count, hour ", hour), axes = TRUE)
    if (nrow(facilities) && all(c("longitude", "latitude") %in% names(facilities))) {
      pts <- tryCatch(terra::project(terra::vect(facilities, geom = c("longitude", "latitude"), crs = "EPSG:4326"), terra::crs(r)), error = function(e) NULL)
      if (!is.null(pts)) terra::points(pts, pch = 21, bg = "#74add1", col = "#333333", cex = 0.8)
    }
    source_id <- if ("source_facility_id" %in% names(selected)) as.character(selected$source_facility_id[i]) else as.character(selected$source_id[i]); source <- facilities[facilities$facility_id == source_id, , drop = FALSE]
    if (nrow(source)) {
      src <- tryCatch(terra::project(terra::vect(source, geom = c("longitude", "latitude"), crs = "EPSG:4326"), terra::crs(r)), error = function(e) NULL)
      if (!is.null(src)) terra::points(src, pch = 17, col = "#d73027", cex = 1.4)
    }
    plotted <- plotted + 1L
  }
}
if (!plotted) cat("No completed saved particle-count rasters are available for plotting.")
```

## Receptor exchange and continuous exposure

```{r}
files <- list.files(file.path(root, "runs"), pattern = "source_receptor_exchange[.]csv$", recursive = TRUE, full.names = TRUE)
exchange <- if (length(files)) do.call(rbind, lapply(files, read_csv)) else data.frame()
if (!nrow(exchange)) {
  cat("Receptor outputs are incomplete or absent. Missing output is not interpreted as zero exposure.")
} else {
  intercept <- if ("intercept" %in% names(exchange)) as.logical(exchange$intercept) else if ("binary_intercept" %in% names(exchange)) as.logical(exchange$binary_intercept) else rep(NA, nrow(exchange))
  n_intercepts <- sum(intercept %in% TRUE)
  cat("Directed source-receptor records:", nrow(exchange), "\n\n")
  cat("Binary intercepts:", n_intercepts, "of", sum(!is.na(intercept)), "\n\n")
  if ("intercept_reason" %in% names(exchange)) {
    cat("Outcome categories (zero modeled exposure, below-threshold exposure, outside-distance screening, and unavailable metrics remain distinct where present):\n\n")
    print(knitr::kable(as.data.frame(table(outcome = exchange$intercept_reason, useNA = "ifany"))))
  }
  continuous <- intersect(c("particle_count_max", "particle_count_cumulative"), names(exchange))
  if (length(continuous)) {
    summaries <- do.call(rbind, lapply(continuous, function(metric) {
      x <- suppressWarnings(as.numeric(exchange[[metric]])); available <- is.finite(x)
      data.frame(metric = metric, available_pairs = sum(available), zero_pairs = sum(available & x == 0), nonzero_pairs = sum(available & x > 0), minimum = if (any(available)) min(x[available]) else NA_real_, median = if (any(available)) stats::median(x[available]) else NA_real_, maximum = if (any(available)) max(x[available]) else NA_real_)
    }))
    cat("\n\nContinuous particle-count exposure summaries:\n\n"); print(knitr::kable(summaries, digits = 3))
  }
  if (n_intercepts == 0L) {
    cat("\n\n**No source–receptor pairs met the configured binary intercept criterion in these runs.**")
  } else if (all(c("source_id", "receptor_id") %in% names(exchange))) {
    cat("\n\nOutgoing intercepts by source:\n\n"); print(knitr::kable(as.data.frame(table(source_id = exchange$source_id[intercept %in% TRUE]))))
    cat("\n\nIncoming intercepts by receptor:\n\n"); print(knitr::kable(as.data.frame(table(receptor_id = exchange$receptor_id[intercept %in% TRUE]))))
    matrix <- as.data.frame.matrix(table(exchange$source_id[intercept %in% TRUE], exchange$receptor_id[intercept %in% TRUE]))
    cat("\n\nPositive-intercept source × receptor matrix:\n\n"); print(knitr::kable(matrix))
    if (all(c("source_longitude", "source_latitude", "receptor_longitude", "receptor_latitude") %in% names(exchange))) {
      plot(facilities$longitude, facilities$latitude, pch = 21, bg = "#74add1", xlab = "Longitude", ylab = "Latitude", main = "Detected source-receptor links")
      hit <- which(intercept %in% TRUE); segments(exchange$source_longitude[hit], exchange$source_latitude[hit], exchange$receptor_longitude[hit], exchange$receptor_latitude[hit], col = "#d73027", lwd = 2)
    }
  }
}
```

## Diagnostic result and key files

Manifest coverage: **`r if (coverage_ok) "PASS" else "FAIL"`**.<br>
Overall diagnostic: **`r diagnostic`**.

Key paths: `r paste(file.path(root, c("inputs/run_manifest.csv", "combined/run_audit.csv", "combined/run_status_summary.csv", "meteorology/meteorology_inventory.csv")), collapse = "; ")`
